Temporal stability in human interaction networks: sector sizes, topological prominence and time activity

### 2015-10-05 paper.tex
arrumada secao dos tipos

\subsection{Types from Erd\"os sectors}\label{sec:pty}
The type of the Erd\"os sector to which a participant belongs 
can be regarded as implying a type to the corresponding participant.
Assigning a type to a participant raises important 
issues about the potential for stigmatization and prejudice. 
In the present case, the type of a participant
changes both along time and as
different networks are considered,
and therefore the potential for prejudice of such
participant typology is attenuated~\cite{adorno}. 
In other words, an individual is a hub in a number of networks and peripheral in other networks, and even within the same network he/she probably changes type along time~\cite{4hubs}.
Moreover, such proposed typology of participants bridges exact and human sciences and may be enriched with concepts from other typologies, such as Meyer-Briggs, Pavlov or the authoritarian types of the F-Scale~\cite{adorno}.

We analyzed the temporal evolution of the networks using visualization tools developed for this research~\cite{rcText,versinus} and inspected the raw data to infer the main characteristics of each type. Our main observations may be summarized as follows:
\begin{itemize}
	\item Core hubs usually have intermittent activity. Very stable activity was found only in smaller communities, which is consistent with the literature~\cite{barabasiEvo}.
	\item Typically, the activity of hubs is trivial: they interact as much as possible, in every occasion with everyone.
The activity of peripheral vertices also follows a simple pattern: they interact very rarely, in very few occasions.
Therefore, intermediary vertices seem responsible for the network structure.
Intermediary vertices may exhibit preferential communication to peripheral, intermediary, or hub vertices; can be marked by stable communication partners; can involve stable or intermittent patterns of activity, to point just a few examples of this greater variety of roles.
	\item Some of the most active participants receive many responses with relative few messages sent, and rarely are top hubs.
These seem as authorities and contrast with participants that respond much more than receive responses.
	\item The most obvious community structure, as observed by a high clustering coefficient, i.e. members know each other often, is found mostly in peripheral and intermediary sectors.
\end{itemize}

With focus on the networks as the whole objects of analysis,
we were able to observe a peculiar correlation pattern 
between the number of threads and the number of participants.
When the number of participants exceeds a threshold,
the number of threads displays a positive correlation with the
number of participants.
This finding is illustrated in Figure~\ref{fig:nmgamma3d} and can also be observed in Table~\ref{tab:genLists}.
Obviously, network types can be derived from such results, which was not attempted here but 
left for future work.

\begin{figure}
   \centering
        \includegraphics[width=\columnwidth]{figs/mpgamma____________}
	\caption{A scatter plot of number of messages (M) versus number of participants (N) versus number of threads ($\Gamma$) for 140 email lists. Highest numbers of threads are found in lists with few participants. The correlation between $N$ and $\Gamma$ is negative for low values of $N$ but positive otherwise. This negative correlation between $N$ and $\Gamma$ can also be observed in Table~\ref{tab:genLists}. For $M=20000$ messages, positive correlation of $N$ and $\Gamma$ is present mostly above 1500 participants. All LAU, LAD, MET lists present smaller networks.}
	\label{fig:nmgamma3d}
\end{figure}

@@

Further work should expand the analysis to include
more types of networks and more metrics.
The data and software needed to reach these results
might also receive dedicated and in-depth
documentation as they enable a greater level of transparency
and work share,
which is adequate to both benchmarking
and the study of systems constituted
by human individuals (see Section~\ref{sec:data}).
The derived typology of hub, intermediary and peripheral participants
was used for semantic web and participatory democracy efforts,
and might be enhanced to yield scientific knowledge~\cite{pnud3,pnud4,pnud5}.
Also, we should pursue the publication of the visualization and sonification
processes used for this research~\cite{versinus,4hubs} and
of the extreme linguistic differences found in each Erd\"os sectors~\cite{rcText}.

@@

\begin{acknowledgments}
	Financial support was obtained from CNPq (140860/2013-4,
	project 870336/1997-5), United Nations Development Program (contract: 2013/000566; project BRA/12/018) and FAPESP. 
	The authors are grateful to the American Jewish Committee for maintaining an online copy of the Adorno book used on the epigraph~\cite{adorno}, to GMANE creators and maintainers, and to the communities of the email lists and other groups used in the analysis, and to the Brazilian Presidency of the Republic for keeping Participabr code and data open.
	We are also grateful to developers and users of Python scientific tools.
\end{acknowledgments}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\nocite{*}
\bibliography{paper}% Produces the bibliography via BibTeX.

\end{document}
%
% ****** End of file aipsamp.tex ******

>>>>

\subsection{Types from Erd\"os sectors}\label{sec:pty}
Assigning a type to a participant raises important 
issues about the scientific cannon for human ypes
and the potential for stigmatization and prejudice. 
The type of the Erd\"os sector to which a participant belongs 
can be regarded as implying a type to the corresponding participant.
In this case, the type of a participant
changes both along time and as
different networks are considered,
and therefore the potential for prejudice of such
participant typology is attenuated~\cite{adorno}. 
In other words, an individual is a hub in a number of networks and peripheral in other networks,
and even within the same network he/she most probably changes type along time~\cite{animacoes}.

The importance of this issue can be grasped by
the consideration of static types derived from quantitative criteria.
For example, in email lists with a small number of participants,
the number of threads has a negative correlation
with the number of participants.
When the number of participants exceeds a threshold,
the number of threads has a positive correlation
with the number of participants.
This finding is illustrated in Figure~\ref{fig:nmgamma3d}
and can also be observed in Table~\ref{tab:genLists}.
The attribution of types to individuals, derived from 
this analysis of the systems, has more potential for prejudice because
the derived participant type is static.
The types, in this latter case, fails to acknowledge that
human individuals are not immutable entities.

Beyond the results discussed above and in previous sections,
our main observations regarding the Erd\"os sectors
and the implicit participant types,
consistent with the literature~\cite{barabasiEvo},
are that
1) hubs and intermediary participants usually have intermittent activity.
Stable activity was found only in smaller communities.
For example, MET list have stable hubs while all LAU, LAD and CPP
present intermittent hubs.
2) Network structure seems to be most influenced by the
activity of intermediary participants as they have less extreme
roles than hubs and peripheral participants and
can therefore connect to the sectors and other participants 
in a more selective and explicit manner.

%Moreover, such typology of participants bridges exact and human sciences and may 
%be enriched with concepts from other typologies,
%such as Meyer-Briggs, Pavlov or the authoritarian types of the F-Scale~\cite{adorno}.

%We analyzed the temporal evolution of the networks
%using visualization
%tools developed for this research~\cite{rcText,versinus}
%and inspected raw data.

%dictated (or revealed) by the
%(e.g. stable or intermittent patterns of activity and preferential communication
%with hubs or periphery)

%of both hubs and peripheral vertices
%have the trivial facets of interacting 
%
%
%
%\begin{itemize}
%	\item Typically, the activity of hubs is trivial: they interact as much as possible, in every occasion with everyone.
%The activity of peripheral vertices also follows a simple pattern: they interact very rarely, in very few occasions.
%Therefore, intermediary vertices seem responsible for the network structure.
%Intermediary vertices may exhibit preferential communication to peripheral, intermediary, or hub vertices; can be marked by stable communication partners; can involve stable or intermittent patterns of activity, to point just a few examples of this greater variety of roles.
%%	\item Some of the most active participants receive many responses with relative few messages sent, and rarely are top hubs.
%%These seem as authorities and contrast with participants that respond much more than receive responses.
%%	\item The most obvious community structure, as observed by a high clustering coefficient, i.e. members know each other often, is found mostly in peripheral and intermediary sectors.
%\end{itemize}

%Within networks as the whole objects of analysis,
%we were able to observe a peculiar correlation pattern 
%between the number of threads and the number of participants.
\begin{figure}
   \centering
        \includegraphics[width=\columnwidth]{figs/mpgamma____________}
	\caption{A scatter plot of number of messages (M) versus number of participants (N) versus number of threads ($\Gamma$) for 140 email lists. Highest numbers of threads are found in lists with few participants. The correlation between $N$ and $\Gamma$ is negative for low values of $N$ but positive otherwise. This negative correlation between $N$ and $\Gamma$ can also be observed in Table~\ref{tab:genLists}. For $M=20000$ messages, positive correlation of $N$ and $\Gamma$ is present mostly above 1500 participants. All LAU, LAD, MET lists present smaller networks.}
	\label{fig:nmgamma3d}
\end{figure}

@@

Further work should expand the analysis to include
more types of networks and more metrics.
The data and software needed to reach these results
might also receive dedicated and in-depth
documentation as they enable a greater level of transparency
and work share,
which is adequate to both benchmarking
and the study of systems constituted
by human individuals (see Section~\ref{sec:data}).
The derived typology of hub, intermediary and peripheral participants
was used for semantic web and participatory democracy efforts,
and might be enhanced to yield scientific knowledge~\cite{opa}.
Also, we should pursue the publication of the visualization and sonification
processes used for this research~\cite{versinus,animacoes} and
of the extreme linguistic differences found in each Erd\"os sectors~\cite{rcText}.

@@

\begin{acknowledgments}
	Financial support was obtained from CNPq (140860/2013-4,
	project 870336/1997-5), United Nations Development Program (contract: 2013/000566; project BRA/12/018) and FAPESP. 
	The authors are grateful to the American Jewish Committee for maintaining an online copy of the Adorno book used on the epigraph~\cite{adorno}, to GMANE creators and maintainers for the public email list data, to the communities of the email lists and other groups used in the analysis, and to the Brazilian Presidency of the Republic for keeping Participabr code and data open.
	We are also grateful to developers and users of Python scientific tools.
\end{acknowledgments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\nocite{*}
\bibliography{paper}% Produces the bibliography via BibTeX.

\end{document}
